\begin{enumerate}
\item
\trigloss[preamble={\textipa{ma\R{}lIn 2 nU\R{}i nU sin:ia mINEnlIt\h{}}}]
{\textipa{ma\R{}lIn} \textipa{2} \textipa{nU\R{}i} \textipa{nU} \textipa{sin:i} -\textipa{a} \textipa{mINEn} -\textipa{lIt\h{}}}
{Marlin and.\textsc{conj} Dory to.\textsc{prep} Sydney -\textsc{sg.adp} travel -\textsc{3.du.prs}}
{Marlin and Dory to Sydney {} travel {}}
{Marlin and Dory are travelling to Sydney.}

\item
\trigloss[preamble={\textipa{nU\R{}i laiA\R{} S:eS:a sAmA 2 ZEnAni\R{} mEa}}]
{\textipa{nU\R{}i} \textipa{lai} -\textipa{A\R} \textipa{S:eS:} -\textipa{A} \textipa{sAm} -\textipa{A} \textipa{2} \textipa{ZEn} -\textipa{A} -\textipa{ni\R} \textipa{mE} -\textipa{a}}
{Dory direction -\textsc{sg.acc} fish -\textsc{sg.adp} dark -\textsc{sg.adp} and.\textsc{conj} mysterious -\textsc{sg.adp} -\textsc{do.post} request -\textsc{3.sg.prs}}
{Dory direction {} fish {} dark {} and mysterious {} {} request {}}
{Dory asks a dark and mysterious fish for directions.}

\item
\trigloss[preamble={\textipa{ma\R{}lIn ZaA\R{} x\oe{}xia sWil S:eS:Ani\R{} sa:i:na}}]
{\textipa{ma\R{}lIn} \textipa{ZaA\R{}} \textipa{x\oe{}xi} -\textipa{a} \textipa{sWil} \textipa{S:eS:} -\textipa{A} -\textipa{ni\R} \textipa{sa:i:n} -\textipa{a}}
{Marlin \textsc{3.sg.acc.pron} reprimand -\textsc{3.sg.pres} while.\textsc{conj} fish -\textsc{sg.adp} -\textsc{do.post} shout -\textsc{3.sg.pres}}
{Marlin her reprimand -s while fish {} to shout -s}
{Marlin reprimands her while she shouts to the fish.}

\item
\trigloss[preamble={\textipa{S:eS: nExmla mIt\h{} ap\h{} S:eS:A\R{} naUnIlA\R{} H2\R{}lIt\h{}}}]
{\textipa{S:eS:} \textipa{nExml} -\textipa{a} \textipa{mIt\h} \textipa{ap\h} \textipa{S:eS:} -\textipa{A\R} \textipa{naUnIl} -\textipa{A\R} \textipa{H2\R} -\textipa{lIt\h}}
{fish grow -\textsc{3.sg.pres} and.\textsc{conj} \textsc{is?} fish -\textsc{sg.acc} whale -\textsc{sg.acc} see -\textsc{3.du.pres}}
{fish grow -s and that fish {} whale {} see {}}
{The fish grows and they see that the fish is a whale.}

\item
\trigloss[preamble={\textipa{nU\R{}i naUnIlAni\R{} naUma}}]
{\textipa{nU\R{}i} \textipa{naUnIl} -\textipa{a} -\textipa{ni\R} \textipa{naUm} -\textipa{a}}
{Dory whale -\textsc{sg.adp} -\textsc{do.post} sing -\textsc{3.sg.prs}}
{Dory whale {} to sing -s}
{Dory sings to the whale.}

\item
\trigloss[preamble={\textipa{mEslAni!ni\R{} laiA\R{} alzini}}]
{\textipa{mE}- \textipa{slani!} -\textipa{ni\R} \textipa{lai} -\textipa{A\R} \textipa{alz} -\textipa{ini}}
{\textsc{imp}- \textsc{1.excl.du.adp.pron} -\textsc{do.post} direction -\textsc{sg.acc} give -\textsc{2.sg.fut}} 
{{} us to direction {} give {}}
{Give us directions.}

\item
\trigloss[preamble={\textipa{naUnIl Us2mIn:a mIt\h ma\R{}lIn Ananak\h{}}}]
{\textipa{naUnIl} \textipa{Us2}- \textipa{mIn:} -\textipa{a} \textipa{mIt\h} \textipa{ma\R{}lIn} \textipa{Ananak\h}}
{whale \textsc{neg}- appear -\textsc{3.sg.pres} and.\textsc{conj} Marlin frightened}
{whale dis- appear -s and Marlin frightened}
{The whale disappears and Marlin is frightened.}

\item
\trigloss[preamble={\textipa{nU\R{}i ZaA\R{} amy\R{}a ap\h{} naUnIl\R{} nEm\R{} SISI\R{} slIx\R{} !amsIml mlAt\h{} ma\R{}lInA\R{} s2!amsIml}}]
{\textipa{nU\R{}i} \textipa{ZaA\R} \textipa{amy\R} -\textipa{a} \textipa{ap\h} \textipa{naUnIl} -\textipa{\R} \textipa{nEm} -\textipa{\R} \textipa{SISI} -\textipa{\R} \textipa{slIx} -\textipa{\R} \textipa{!ams} -\textipa{Iml} \textipa{mlAt\h} \textipa{ma\R{}lIn} -\textipa{A\R} \textipa{s2}- \textipa{!ams} -\textipa{Iml}}
{Dory \textsc{3.sg.acc.pron} tells -\textsc{3.sg.pres} \textsc{is} whale -\textsc{pl.acc} all -\textsc{pl.acc} krill -\textsc{pl.acc} some -\textsc{pl.acc} eat -\textsc{3.pl.fut} and.\textsc{conj} Marlin \textsc{neg}- eat -\textsc{3.pl.fut}}
{Dory him tell -s that whale -s all {} krill {} some {} eat {} and Marlin not eat {}}
{Dory tells him that whales eat krill, not Marlins.}

\item
\trigloss[preamble={\textipa{ZasWs ZEn:im\R{}a SISI\R{} aila\R{} H2\R{}lIt\h}}]
{\textipa{ZaA} -\textipa{sWs} \textipa{ZEn:i} -\textipa{m\R{}a} \textipa{SISI} -\textipa{\R} \textipa{aila} -\textipa{\R} \textipa{H2\R} -\textipa{lIt\h}}
{\textsc{3.sg.nom.pron} -at.\textsc{post} \textsc{3.du.adp} -near.\textsc{post} krill -\textsc{pl.acc} many -\textsc{pl.acc} see -\textsc{3.du.pres}}
{it at them near krill {} many {} see {}}
{Suddenly they see a lot of krill near them.}

\item
\trigloss[preamble={\textipa{mEAnaNy\R{}e 2mzini}}]
{\textipa{mE}- \textipa{AnaN} -\textipa{y\R{}e} \textipa{2mz} -\textipa{ini}}
{\textsc{imp}- fear -for-the-sake-of.\textsc{post} pause -\textsc{2.sg.fut}}
{{} fear for pause {}}
{Pause for suspense.}

\item
\trigloss[preamble={\textipa{naUnIl ZEn:i\R{} Hama}}]
{\textipa{naUnIl} \textipa{Zen:i\R} \textipa{Ham} -\textipa{a}}
{whale \textsc{3.du.acc.pron} swallow-whole -\textsc{3.sg.pres}}
{whale them swallow-whole -s}
{The whale swallows them whole.}

\item
\trigloss[preamble={\textipa{ma\R{}lIn naUnIl\R{} fIlminIt\h{} m\R{}SAmIa EmImla}}]
{\textipa{ma\R{}lIn} \textipa{naUnIl} -\textipa{\R} \textipa{fIlmin} -\textipa{IT\h} \textipa{m\R{}sami} -\textipa{a} \textipa{EmIml} -\textipa{a}}
{Marlin whale -\textsc{sg.acc} baleen -\textsc{sg.poss} frantic -\textsc{3.sg.pres} attack -\textsc{3.sg.pres}}
{Marlin whale {} baleen {} frantic -ally attack -s}
{Marlin frantically attacks the whale's baleen.}

\item
\trigloss[preamble={\textipa{ZasWs H2\R{}a ap\h{} naUnIlA\R{} eNlImANIt\h{} w\R{}WIt\h{} min:a}}]
{\textipa{ZaA} -\textipa{sWs} \textipa{h2\R} -\textipa{a} \textipa{ap\h} \textipa{naUnIl} -\textipa{A\R} \textipa{eNlImak\h} -\textipa{It\h} \textipa{w\R{}W} -\textipa{It\h} \textipa{min:} -\textipa{a}}
{\textsc{3.sg.nom.} -at.\textsc{post} see -\textipa{3.sg.pres} \textsc{is?} whale -\textsc{sg.acc} mouth -\textsc{sg.poss} water-surface -\textsc{sg.poss} approach -\textsc{3.sg.pres}}
{it at see {} that whale {} mouth {} water-surface{} approach -ing}
{Suddenly they see that the surface of the water of the whale's mouth is approaching.}

\item
\trigloss[preamble={\textipa{naUnIl ZaA\R{} IlAlAIt\h{} lENa}}]
{\textipa{naUnIl} \textipa{ZaA\R} \textipa{IlAlA} -\textipa{It\h} \textipa{lek\h} -\textipa{a}}
{whale \textsc{3.sg.acc.pron} tongue -\textsc{sg.poss} flap -\textipa{3.sg.pres}}
{whale it tongue {} flap -s}
{The whale flaps its tongue.}

\item
\trigloss[preamble={\textipa{IlAlA\R{} N\R{}ImlIt\h{} sWil nU\R{}i ma\R{}lInA\R{} saTa ap\h{} ZaAsm \R{}iUm:ia mIt\h{} s2N\R{}im:a}}]
{\textipa{IlAlA} -\textipa{\R} \textipa{N\R{}Im} -\textipa{lIt\h} \textipa{sWil} \textipa{nU\R{}i} \textipa{ma\R{}lIn} -\textipa{A\R} \textipa{saT} -\textipa{a} \textipa{ap\h} \textipa{ZaA} -\textipa{sm} \textipa{\R{}iUm:i} -\textipa{a} \textipa{mIt\h} \textipa{s2}- \textipa{N\R{}Im:} -\textipa{a}}
{tongue -\textsc{sg.acc} hold-onto -\textsc{3.du.pres} while Dory Marlin -\textsc{sg.acc} beg -\textipa{3.sg.pres} \textsc{is} \textsc{3.sg.adp.pron} -\textsc{accom.post} rise -\textipa{3.sg.pres} and.\textsc{conj} \textsc{neg} hold-onto -\textsc{3.sg.pres}}
{tongue {} hold-onto {} while Dory Marlin {} beg -s to her with rise {} and not hold-onto {}}
{They hold onto the tongue while Dory begs Marlin to trust her and let go.}

\item
\trigloss[preamble={\textipa{ma\R{}lIn nU\R{}iA\R{} x\oe{}xia mlAt\h{} ZEm nax\R{}A ZaA\R{}lI H2\R{}a}}]
{\textipa{ma\R{}lIn} \textipa{nU\R{}i} -\textipa{A\R} \textipa{x\oe{}xi} -\textipa{a} \textipa{mlAt\h} \textipa{ZEm} -$\varnothing$ \textipa{nax\R} -\textipa{A} \textipa{ZaA} -\textipa{\R{}lI} \textipa{h2\R} -\textipa{a}}
{Marlin Dory -\textsc{sg.acc} scold -\textsc{3.sg.pres} and.\textsc{conj} light-ocean -\textsc{sg.adp} huge -\textsc{sg.adp} \textsc{3.sg.adp.pron} -below.\textsc{post} see -\textsc{3.sg.pres}}
{Marlin Dory {} scold -s and \textipa{ZEm} {} huge {} him below see -s}
{Marlin scolds Dory and sees a huge \textipa{ZEm} below him.}

\item
\trigloss[preamble={\textipa{mEAnaNy\R{}e aN2mzini}}]
{\textipa{mE}- \textipa{AnaN} -\textipa{y\R{}e} \textipa{aN}- \textipa{2mz} -\textipa{ini}}
{\textsc{imp}- fear -for-the-sake-of.\textsc{post} two- pause -\textsc{2.sg.fut}}
{{} fear for again pause {}}
{Pause again for suspense.}

\item
\trigloss[preamble={\textipa{s2N\R{}Im:a}}]
{\textipa{s2}- \textipa{N\R{}Im:} -\textipa{a}}
{\textsc{neg}- hold-onto -\textsc{3.sg.pres}}
{not hold-onto -s}
{He lets go.}

\item
\trigloss[preamble={\textipa{naUnIl ZEn:i\R{} i:aefa Hama mlAt\h{} sin:iasei naUzia}}]
{\textipa{naUnIl} \textipa{ZEn:i\R} \textipa{i:aef} -\textipa{a} \textipa{Ham} -\textipa{a} \textipa{mlAt\h} \textipa{sin:i} -\textipa{sei} \textipa{naUzi} -\textipa{a}}
{whale \textsc{3.du.acc.pron} quick -\textsc{3.sg.pres} swallow-whole -\textsc{3.sg.pres} and.\textsc{conj} Sydney -into.\textipa{post} exhale -\textsc{3.sg.pres}}
{whale them quick -ly swallow-whole -s and Sydney into blow -s}
{The whale swallows them quickly and and blows them into Sydney.}

\end{enumerate}
